---
title: "R Notes; Math Camp 1"
format: pdf
editor: visual
---

```{r setup, include = FALSE}

library(tidyr);
library(tidyverse);
library(dplyr);
library(ggplot2);
library(tinytex);

setwd("/Users/oneone/Library/Mobile Documents/com~apple~CloudDocs/MIT/Fall 2025/Math Camp 1/Math Camp 1/Notes")
```

# Day 1

## Objects

```{r}

x <- 5 # create a variable, x, with value, 5
y <- c(1:10) # create a vector with numbers 1 through 10
z <- matrix(c(1:10), nrow = 5) # create a matrix with 1:10 across 5 rows
```

## Data Frames

```{r}

df <- data.frame(
  name = c("Emily", "Andrea"),
  age = c(26, 33)
) # create a data frame with the name and age of yourself and the person to your right

mean(y)
```

### Loading and Viewing Datasets

```{r}

sdi.dataset <- read_csv("/Users/oneone/Library/Mobile Documents/com~apple~CloudDocs/MIT/Fall 2025/Math Camp 1/Math Camp 1/Datasets/SDI_2.0.csv")

summary(sdi.dataset)
head(sdi.dataset)

```

### Cleaning Datasets

```{r}

clean.SDI.datset <- sdi.dataset %>%
  select("state", "year", "democracy_mcmc")  %>% # select state, year, and democracy_mcmc variables
  dplyr::filter(year >= 2010) # filter observations to years 2010-2023
```

## Functions

### Basics

```{r}

mean(y) # finds the average value of vector, y
mean(df$age) # finds the average value of the variable, age in data frame, df
```

### Creating a Function

```{r}

my.mean <- function(x) {
  sum(x)/length(x)
} # builds a function to measure the mean of object, x

my.mean(df$age)
```

## Packages

### Important packages for Quant 1:

-   tidyverse

-   ggplot2

-   dplyr

-   kableExtra/gt - helpful for creating tables for presentation format

-   modelsummary/stargazer - helpful for regression tables

-   marginaleeffects - helpful for visualizing predictions from regressions

-   sandwhich - helpful for special standard errors for regression
